Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-04-20 WholePlotTrajectories.Rmd
html

To counter the remaining taxonomic uncertainty plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
The analysis of taxonomic and functional composition of plots along time was visualized in a two-dimensional ordination space based on non-metric dimensional scaling of succesive floristic or functional inventories. 
Plots trajectories along time was reported comparatively to the inventories in 1989, 5 years after disturbance, which corresponded to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees). 
The inventories dissimilarity compared to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
The functional trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004]. 
Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were  therefore reported as the proportion of each class in the inventories.

@@

Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
Trajectories of taxonomic and functional composition after disturbance were examined through the ordination of successive inventories from 1989 (5 years after disturbance) in the two-dimensional space from NMDS analysis based on flora inventories and the corresponding functional composition. 
Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity (data not shown).

Both taxonomic and functional composition were substantially affected by disturbance, and the maximum dissimilarity to 1984 reference inventory composition was positively correlated with the disturbance intensity.
Until 30 years after disturbance, all disturbed plots taxonomic composition remained significantly dissimilar to that of the 1989 reference inventory composition (Figure \@ref(fig:NMDSplans)).
All plots, though, displayed a unimodal trajectory with a the return toward the taxonomic and functional composition of the 1989 reference inventory suggesting a shift towards a cyclic regime (Figure \@ref(fig:NMDSdivergence)). 
For taxonomic composition, the maximum dissimilarity to 1984 inventory composition was reach around 20 years after disturbance for the T2 and T3 plots and for one T1 plot while control and other T1 plots continued to increase (Figure \@ref(fig:NMDSdivergence).a). 
For functional composition, the maximum dissimilarity to 1984 inventory composition was reach from 15 to 20 years after disturbance for the T2 and T3 plots while T1 and control plots continued to increase (Figure \@ref(fig:NMDSdivergence).b). 
The functional composition distancing from the reference inventory was also positively correlated with disturbance intensity and similarly stabilized or reduced (for 5 of T1, T2 and T3 plots) from 20 years after disturbance (Figure \@ref(fig:NMDSdivergence)).

The coordinates of the functional traits were measured in the two-dimensional ordination space that mapped plots evolution along time (graphs not showed) revealed a trajectory of disturbed plots toward acquisitive functional strategies (from high WD to high SLA and chlorophyll content).

@@

```{r NMDSdivergence, out.width = '80%', fig.asp = 1.2/2, echo=FALSE,fig.cap="Trajectories of the distance to initial condition of the 30 sampled years in the two-dimensional space from the NMDS of v taxonomic composition at genus level and (ref:Boldb) functional composition. Distance are abundance-based Bary-Curtis metric. Line colors represent the disturbance treatment (green for control, blue for T1,orange for T2 and red for T3). The 0.025 and 0.975 percentile correspond to the variance observed for 50 iteration of the taxonomic uncertainty propagation and functional trait filling processes.", fig.env="figure*"}

TaxoDist(MatrepTaxo)
FunDist(MatrepFun)

@@

For all plots the trajectories of the community weighted means (CWM) were drawn for the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) (Figure \@ref(fig:CWM)).

To compensate the intrinsinc difference among plots the trajectories drawn correspond to the difference in value between the reference inventory in 1984 (5 year after disturbance) and the successive years inventoried.

@@

For undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the initial values 5 years after disturbance.
After disturbance, the richness increased after low disturbance intensity, with a maximum increase of 14 botanical genuses (for plot 3 after treatment 2), but followed a unimodal decrease with a return to initial values after intense disturbance.
The taxonomic evenness (Shannon and Simpson diversities), however, significantly increased after all disturbance regime.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.92$, and $\rho_{spearman}^{Simpson}=0.97$).
Only two T3 plots, plots 8 and 12, remained increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.

>>>>

To counter the remaining taxonomic uncertainty plots taxonomic composition and diversity were analysed at the genus level, *i.e.* referring to the genus of observed or trialed botanical names.
Variations in plots taxonomic and functional composition after disturbance was visualized by their trajectories in a two-dimensional ordination space over 30 years.
Two NMDS were conducted, either from taxonomic flora inventories or from plots functional composition based on the 8 leaf, stem and life history traits. 
Plots trajectories along time was reported comparatively to the inventories in 1989, 5 years after disturbance, which corresponded to first inventory with a sufficient degree of uncertainty (<30% of undetermined trees). 
The inventories dissimilarity compared to the reference 1989 inventory was reported using occurrence-based (Jaccard) and abundance-based (Bray-Curtis) similarity measures. 
The trajectory of inventories along time was visualized with the euclidean distance in the two-dimensional ordination space to the 1989 inventory.
The trajectories of the leaf and stem and life traits were also visualized with the community weighted means (CWM), representing the average trait value in a community weighted by relative abundance of the species carrying each value [@Diaz2007;@Garnier2004]. 
To compensate the intrinsinc difference among plots the trajectories corresponded to the differences along time with the reference inventory in 1984 (5 year after disturbance).
Species seed mass corresponded to 5 classes of increasing mass, seed mass trajectories were  therefore reported as the proportion of each class in the inventories.

@@

Over time, 828388 trees and 591 botanical species pertaining to 223 genus and 64 botanical families were recorded. 
Taxonomic and functional plots trajectories were examined through the ordination in a two dimensional ordiantion space of successive inventories from 1989 (5 years after disturbance) to 2015 (31 years after disturbance).
Classifications were performed using either abundance-based Bray-Curtis (Figure \@ref(fig:NMDSplans)) or incidence-based Jaccard dissimilarity, this last having given similar results (data not shown).

Both taxonomic and functional composition of disturbed plots followed consistent trajectories over time, translating compositional changes while control plots remained stable.
The addition of functional traits in the two dimensional space revealed a shift of plots composition towards more acquisitive functional strategies (disturbed plots changed from high average WD to high average SLA and chlorophyll content).
The dissimilarity of the successive inventories for disturbed plots compared to their respective reference inventory in 1989 followed unimodal trajectories (AppendixI, Figure A2).
The maximum dissimilarity of plot strajectories was positively correlated to the disturbance intensity for taxonomic ($\rho_{spearman}^{taxonomic}=0.91$) as well as functional composition ($\rho_{spearman}^{taxonomic}=0.96$). 
The time at maximum dissimilarity for taxonomic composition was reached around 26 years after disturbance while it was 22 years for functional composition. 
All trajectories returned towards the initial composition and translated cyclic compositional changes as visualized in figure (Figure \@ref(fig:NMDSplans)).

@@

@@

The changes observed in plots functional composition went hand to hand with consistent trajectories of the 8 functional and life history traits (Leaf thickness, chlorophyll content, toughness and specific area,wood specific gravity and barck thickness and seed mass and maximum adult height) visaualized with plots community weighted means (CWM) (Figure \@ref(fig:CWM)).

@@

For undisturbed plots the Richness, Shannon and Simpson diversity remained comparable to the initial values 5 years after disturbance.
After disturbance, the richness increased after low disturbance intensity, with a maximum increase of 14 botanical genuses (for plot 3 after treatment 2), but followed a unimodal decrease with a return to initial values after intense disturbance.
The taxonomic evenness (Shannon and Simpson diversities), however, significantly increased after all disturbance regime.
The evenness followed a unimodal trajectory with a just beginning return towards initial values and a maximum reached around 20 years and positively correlated to the disturbance intensity ($\rho_{spearman}^{Shannon}=0.86$, and $\rho_{spearman}^{Simpson}=0.89$).
Only two T3 plots, plots 8 and 12, remained increasing 30 years after disturbance \@ref(fig:DivTaxo)), suggesting a similar be much delayed unimodal trajectory.
